---
title: "Tutorial on SGD"
author: "Hon Hwang and Alan Malecki"
date: "2016 August"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Stochastic Gradient Descent (SGD) 

- Used heavily in computing
- *Iterative*
    + Step by step
- *Optimisation*
    + Finding min/max of a function. Finding roots.
- *Algorithm*
    + Implementable

## General Form of Iterative Optimisation

<center>$\theta_{k+1} = \theta_{k} - \alpha g(\theta)$</center>

- $k \in {1, 2, 3, \ldots, L}$
    + $L$ is the number of iterations
- $\theta$ can be one parameter or a vector
- $\alpha$ is called *step size*, *learning rate*, or *gain*
    + Can be a scalar or a matrix (e.g., Newton-Ralphson)
- $g(\theta)$ is the gradient function
    + 1st derivative of:
        * *Objective Function*; or
        * *Loss Function*

## Intutively…

## Iris Data

We'll use the `iris` dataset

```{r iris-data, echo = FALSE}
data(iris)
str(iris)
```

## Iris Data - Variables of Interest

```{r iris-lm-vars, ech = FALSE}
library(ggplot2)

qplot(x = Petal.Length, y = Petal.Width, data = iris,
      xlab = "Petal.Length", ylab = "Petal.Width")
```

## `lm()` Results


## SGD Implementation

- Least Mean Square Loss function
    + Equation here
    
- Gradient for 2 parameters
    + Equation here

- General Gradient Equation

## Gradient Descent v SGD

- Instanteous Gradient

## SGD Implementation

```{r sgd-lm, echo = TRUE}
```
