---
title: "Tutorial on SGD"
author: "Hon Hwang and Alan Malecki"
date: "2016 August"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Stochastic Gradient Descent (SGD) 

- Used heavily in computing
- *Iterative*
    + Step by step
- *Optimisation*
    + Finding min/max of a function
    + Finding roots of equations
- *Algorithm*

## General Form of Iterative Optimisation

<center>$\boldsymbol{\theta}_{k+1} = \boldsymbol{\theta}_{k} - \alpha g(\boldsymbol{\theta})$</center>

- $k \in {1, 2, 3, \ldots, K}$
    + $K$ is the number of iterations
- $\boldsymbol{\theta}$ is the vector of parameters to estimate, or just a
  single parameter
- $\alpha$ is called *step size*, *learning rate*, or *gain*
    + Can be a scalar or a matrix (e.g., Newton-Ralphson)
- $g(\boldsymbol{\theta})$ is the gradient function
    + $g(\boldsymbol{\theta})$ = $\triangledown L(\boldsymbol{\theta})$ 
    + 1st partial derivative of L($\boldsymbol{\theta}$):
        * *Objective Function*; or
        * *Loss Function*

## Notations

- Depending on the context, notations may differ

In Equation (2) of Bottou (2010):

- $k$, the current iteration number is represented as $w$
- The step size $\alpha$ is $\gamma$

- Just keep in mind the general form

## Intuitively…

- Attempting to find the valley (lowest point) of a surface. E.g.,
  The UTS Green area.
- $\alpha$ (step size/learning rate/gain)
     + How big is your next stride
- $g(\boldsymbol{\theta})$, or $\triangledown L(\boldsymbol{\theta})$
     + The *direction* to take the next step
- Since you want to find the valley, the ideal direction is the
  *steepest* path

## Example - Iris Data

We'll use the `iris` dataset

```{r iris-data, echo = FALSE}
data(iris)
str(iris)
```

## Iris Data - Variables of Interest

```{r iris-lm-vars, echo = FALSE}
library(ggplot2)

qplot(x = Petal.Length, y = Petal.Width, data = iris,
      xlab = "Petal.Length", ylab = "Petal.Width")
```

## `lm()` Results

```{r iris-lm, echo = FALSE}
iris_lm_fit <- lm(Petal.Width ~ Petal.Length, data = iris)
```

```{r iris-lmfit}
summary(iris_lm_fit)
```

## SGD Implementation

```{r sgd-lm, echo = TRUE}
```

## Limitations of SGD

- Approximate to true parameters
- May not converge
- Local minimum
- Bottou (2010) refers to Robbins and Siegmund (1971), re-published in 1985 as 
  the source for the theory on convergence

## Types of optimisation

Recall the general form of iterative optimisation:

<center>$\boldsymbol{\theta}_{k+1} = \boldsymbol{\theta}_{k} - \alpha g(\boldsymbol{\theta})$</center>
<br />

- Algorithms differ in how they specify $\alpha$, the step size, and
  $g(\boldsymbol{\theta})$, the gradient function

## Types of optimisation

- Different $\alpha$:
      + In Newton–Raphson, it's the inverse of, the matrix containing the 2nd partial derivatives of the loss function
          * Hessian
      + In SGD, it's a scalar value
- Different $g(\boldsymbol{\theta})$:
      + Sum over all the observations
      + One obsercation at a time
- *Restricted optimisaton* is when a constraint is placed on the 
  $\boldsymbol{\theta}$ iteself. E.g., Lasso

## Implementations

- *Back propagation* in Neural Networks
    + Use of chain rule of differentiation when calculating the gradient. Minnar (2015)
- Parallelised implementation
    + Dean, et al. (2012). Google's *DistBelief*. A *asynchornous* SGD implementation.
    + Chen, et al. (2016). A *synchornous* SGD implementation.
- Apache Manhout, Apache Spark - non-parallel but applicable to large data sets.                                      
    
## References

Bottou, L. 2010, *Large-scale machine learning with stochastic gradient descent*,
Proceedings of COMPSTAT’2010, pp. 177–86.

Chen, J., Monga, R., Bengio, S. & Jozefowicz, R. 2016,
*Revisiting Distributed Synchronous SGD*, ICLR, p. 5.

Dean, J., Corrado, G.S., Monga, R., Chen, K., Devin, M., Le, Q. V, Mao, M.Z.,
Ranzato, M.A., Senior, A., Tucker, P., Yang, K. & Ng, A.Y. 2012,
*Large Scale Distributed Deep Networks*,
NIPS 2012: Neural Information Processing Systems, pp. 1–11.

Minnaar, A. 2015, *Deep Learning Basics: Neural Networks, Backpropagation and Stochastic Gradient Descent*,
viewed 16 August 2016, 
[](http://alexminnaar.com/deep-learning-basics-neural-networks-backpropagation-and-stochastic-gradient-descent.html)

## References

Robbins, H. & Siegmund, D. 1985,
*A Convergence Theorem for Non Negative Almost Supermartingales and Some Applications*,
in T.L. Lai & D. Siegmund (eds),Herbert Robbins Selected Papers,
Springer New York, New York, NY, pp. 111–35.

Spall, James C., 2003. *Introduction to stochastic search and optimization estimation, simulation, and control*,
Wiley-Interscience, Hoboken, N.J.
